Smart Model Elimination for Efficient Automated Machine Learning in Healthcare

### 2024-06-25 paper/paper.tex
improve abstract and summary

\begin{abstract}
  Automated Machine Learning or AutoML has emerged as a popular field of research. We perform a literature review of existing AutoML papers and conduct a survey on five popular AutoML frameworks. Many of these frameworks optimize a form of model selection, but they all require every model to be run and evaluated. We propose a novel framework, SMEML (Smart Model Elimination Machine Learning), that automatically eliminates models that are unlikely to be performant. SMEML demonstrates the ability to generate a model with near equal accuracy to a "dumb" brute-force framework, but in a fraction of the time. We also believe this innovation is particularly relevant to machine learning in healthcare because of the necessity for efficient diesease prediction. 
\end{abstract}

@@

Several observations were made from running these datasets. First, accuracy did not change for 4 of the 6 benchmarks. The other two benchmarks on eye states and diabetes prediction represented a loss of accuracy for SMEML by 0.13\% and 1.95\%, respectively. Training time, however, favored SMEML in every trial. SMEML outperformed the dummy framework by 173.96 seconds in its worst performance on the lung cancer trial and by 383.94 seconds in its best performance on the eye state dataset. Relative to the dummy's performance, a percent decrease in time to run for SMEML of at least 46.78\% could be seen, with a decrease of up to 69.22\% observerd as well. SMEML's worst performance by percent decrease was also the performance in which it returned perfect accuracy. 
\section{Discussion}
\subsection{Summary}
SMEML manages to return comparable accuracy results to its dumb couterpart while more than splitting the run time in half in the majority of benchmarks. Its performance is not limited by the size of the dataset, and it returned its first, third, and fourth overall best run times for SMEML relative to the dumb mode on the 'large' dataset benchmarks (those exceeding 1000 rows). This is important in third world medicine where time is of the essence and there is a constant influx of new patient data. A practical demonsration of this software's real-world application can be found on our SmartDx website [23]. Source code can be found at [24].
\subsection{Limitations}
Our implementation of SMEML has limitations. These include a small training set due to lack of available datasets, finite computing power, time constraints, and a lack of access to paticular scientific literature. Our framework only improves on the model selection process and does not address the various other aspects of AutoML. Therefore, SMEML is not comparable to other full-featured AutoML frameworks.
\subsection{Future Work}
Future work will include increasing the size of the training set, implementing a fully fledged AutoML framework, and conducting a detailed experiment comparing SMEML to other popular AutoML frameworks. Furthermore, we will examine other types of machine learning such as regression and multiclass-classification.

@@

[21] fedesoriano, "Stroke Prediction Dataset" (dataset). \url{www.kaggle.com/fedesoriano/stroke-prediction-dataset}.

>>>>

\begin{abstract}
  Automated Machine Learning (AutoML) has emerged as a popular field of research. We present a literature review of existing AutoML papers and conduct a survey on five popular AutoML frameworks. Typically, these frameworks engage in model selection by requiring every model to be run and assessed, a process both time-intensive and computationally expensive. In response, we propose a novel framework, Smart Model Elimination Machine Learning (SMEML), that strategically eliminates models that are unlikely to yield high accuracy. SMEML demonstrates the ability to achieve comparable accuracy to a traditional brute-force approach while significantly reducing the time required. We also believe this innovation is particularly beneficial to machine learning in healthcare, where efficient and accurate disease prediction is crucial. 
\end{abstract}

@@

Several observations were made from running these datasets. First, accuracy did not change for 4 of the 6 benchmarks. The other two benchmarks on eye states and diabetes prediction represented a loss of accuracy for SMEML by 0.13\% and 1.95\%, respectively. Training time, however, favored SMEML in every trial. SMEML outperformed the dummy framework by 173.96 seconds in its worst performance on the lung cancer trial and by 383.94 seconds in its best performance on the eye state dataset. Relative to the dummy's performance, a percent decrease in time to run for SMEML of at least 46.78\% could be seen, with a decrease of up to 69.22\% observerd as well. SMEML's worst performance by percent decrease was also the performance in which it returned perfect accuracy. 
\section{Discussion}
\subsection{Summary}
SMEML consistently achieves comparable accuracy to its dumb counterpart while significantly reducing the time required to train the model. Notably, SMEML's performance is not limited by the size of the dataset, achieving its best runtimes--first, third, and fourth fastest overall--on 'large' datasets (those exceeding 1000 rows). This efficiency is important in third world medicine where there is a constant influx of new patient data that must be processed quickly. A practical demonstration of this software's real-world application can be found on our SmartDx website [23]. Source code can be found at [24]. 
\subsection{Limitations}
Our implementation of SMEML has limitations. These include a small training set due to lack of available datasets, finite computing power, time constraints, and a lack of access to paticular scientific literature. Our framework only improves on the model selection process and does not address the various other aspects of AutoML. Therefore, SMEML is not comparable to other full-featured AutoML frameworks.
\subsection{Future Work}
Future work will include increasing the size of the training set, implementing a fully fledged AutoML framework, and conducting a detailed experiment comparing SMEML to other popular AutoML frameworks. Furthermore, we will examine other types of machine learning such as regression and multiclass-classification.

@@

[21] fedesoriano, "Stroke Prediction Dataset" (dataset). \url{https://kaggle.com/fedesoriano/stroke-prediction-dataset}.
